---
title: "An Introduction to rtweet: <br> Getting & Analyzing Data <br> from the Twitter API"
subtitle: "SatRdaysDC"
author: "Matthew Hendrickson"
date: "2020/03/28"
output:
  ioslides_presentation:
    widescreen: true
    incremental: true
    fig_width: 9
    fig_height: 3
    df_print: paged
#logo: images/rtweet.png # Shows on all slides
---

```{r setup, include = FALSE, echo = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE, comment = "#>", collapse = TRUE)
```

## Topics
1. About Me
2. rtweet
3. The Twitter API
4. `search_tweets()`
5. `get_trends()`
6. `stream_tweets()`
7. Social Network Analysis - `get_friends() get_followers()`
8. `get_timelines()`
9. `get_favorites()`
10. But wait! There's more!



## About Me
<div style="float: left; width: 50%;">
<img src = "images/headshot.png" width="450" height="450">
</div>

<div style="float: left; width: 50%;">
- Social Scientist by Training
     - Psychology & Music `%>%`
     - More Psychlogy `%>%`
     - Law & Policy
- Higher Education Analyst by Trade
- R User by Stumbling
     - Excel `%>%`
     - SPSS GUI `%>%`
     - SPSS Syntax `%>%`
     - SQL `%>%`
     - R
</div>



## rtweet
<div style="float: left; width: 50%;">
<img src = "images/rtweet.png" width="450" height="450">
</div>

<div style="float: left; width: 50%;">
- [rtweet Vignette - https://rtweet.info/](https://rtweet.info/)
- Developed & maintained by [\@kearneymw](https://twitter.com/kearneymw)
- An implementation of calls designed to collect and organize Twitter 
data via Twitter's REST and stream Application Program Interfaces (API), 
which can be found at the following URL:  
<https://developer.twitter.com/en/docs>.
</div>


## The Twitter API
- No longer need a Twitter developer account
     - Walthrough [available](https://rtweet.info/articles/auth.html) for developer accounts
- You _*DO*_ need a Twitter account
- Authorize API by calling an `rtweet` function like `search_tweets()`
- Twitter API limits 18,000 results / 15 min
     - `retryonratelimit = TRUE`
- Must be used in accordance with [Twitter's developer terms](https://developer.twitter.com/en/docs/tweets/data-dictionary/overview/tweet-object)



```{r environment setup, echo = FALSE, message = FALSE, warning = FALSE, eval = TRUE}
library("rtweet")    # Needed to get data from Twitter
library("httpuv")    # Needed to use rtweet without developer account
library("tidyverse") # Data manipulation and visualization
library("scales")    # Customize axes
library("UpSetR")    # To create upset plot
library("igraph")    # For SNA
```



## Tweeting with rtweet
```{r tweeeting, echo = TRUE, message = FALSE, warning = FALSE, eval = FALSE}
post_tweet(
  status = "Hey @Satrdays_DC, did you know you can 
            post a Tweet directly from #rstats?",
  media = "images/no_way.gif"
)
```

<center><img src = "images/no_way.gif" width="500" height="279"></center>
<center><font size="2">[https://memecrunch.com/meme/C0QML/no-way](https://memecrunch.com/meme/C0QML/no-way)</font></center>




## `search_tweets()` {.smaller}
```{r search tweets, echo = TRUE, message = FALSE, warning = FALSE, eval = TRUE}
search_tweets("#rladies", n = 5000, include_rts = FALSE) %>% 
ts_plot("days", trim = 1L) +
  geom_point() + theme_minimal() +
  theme(plot.title = element_text(face = "bold")) +
  labs(x = NULL,  y = NULL,
    title = "Frequency of #rladies Twitter Hashtag Usage",
    caption = "\nSource: Data collected via rtweet - graphic by @mjhendrickson")
```



## `get_trends()` {.smaller}
``` {r trends, echo = TRUE, message = FALSE, warning = FALSE, eval = TRUE}
get_trends("washington") %>% # trends_available() to see locales
  top_n(3, tweet_volume) %>% # Pull top 3 trends by volume
ggplot() +
  geom_bar(mapping = aes(x = reorder(trend, -tweet_volume), y = tweet_volume), stat = "identity") +
  scale_y_continuous(labels = comma) +
  theme(axis.title.x = element_blank()) +
  theme(axis.title.y = element_blank()) +
  labs(title = "Washington DC Twitter Trends",
       caption = "\nSource: Data collected via rtweet - graphic by @mjhendrickson")
```



## `stream_tweets()` {.smaller}
```{r stream tweets, echo = TRUE, message = FALSE, warning = FALSE, eval = FALSE}
rstats <- stream_tweets("rstats", timeout = 60 * 60) # = 60 min
rstats %>%
ts_plot("mins", trim = 1L) +
  geom_point() + theme_minimal() +
  theme(plot.title = element_text(face = "bold")) +
  labs(x = NULL,  y = NULL,
    title = "Frequency of #rstats Twitter Mentions",
    caption = "\nSource: Data collected via rtweet - graphic by @mjhendrickson")
```

<center><img src = "images/plot_rstats_stream.png" width="500" height="279"></center>



## Social Network Analysis {.smaller}
``` {r sna, echo = TRUE, message = FALSE, warning = FALSE, eval = FALSE}
# Get list of friends
friends        <- get_friends(c("mjhendrickson", "kearneymw", "LittleMissData", "AnnaHenschel"))

# Get screen names for friends
friends_names  <- lookup_users(friends$user_id)

# Add screen names to friends data
friends_list   <- left_join(friends, friends_names)
friends_list   <- friends_list[, c("user", "screen_name")]

# Count of accounts followed by users
friends_data   <- table(friends_list$screen_name)

# User overlap
friends_trim   <- subset(friends_list, screen_name %in% names(friends_data[friends_data >2L]))

# Create matrix
friends_matrix <- as.matrix(friends_trim)

# Plot SNA
plot(graph_from_edgelist(friends_matrix), vertex.color = "grey", vertex.label.color = "black", 
     vertex.label.cex = .75, edge.curved = .25, edge.color = "grey20", arrow.size = 5, arrow.width = .25)
```



## Social Network Analysis Output
<center><img src = "images/plot_sna.png"></center>



## `get_timelines()` {.smaller}
``` {r timelines, echo = TRUE, message = FALSE, warning = FALSE, eval = TRUE}
get_timelines(c("dataandme", "thomas_mock", "CMastication"), n = 5000) %>%
  filter(created_at > "2020-01-01") %>% group_by(screen_name) %>% 
ts_plot("days", trim = 1L) +
  geom_point() + theme_minimal() +
  theme(legend.position = "bottom",
    plot.title = element_text(face = "bold")) +
  labs(x = NULL, y = NULL,
    title = "Frequency of Tweets by Cool #rstats Twitter Folks",
    caption = "\nSource: Data collected via rtweet - graphic by @mjhendrickson")
```



## `get_favorites()` {.smaller}
``` {r favorites, echo = TRUE, message = FALSE, warning = FALSE, eval = TRUE}
hadley_favs <- get_favorites('hadleywickham', n = 1000)
hadley_fav_count <- hadley_favs %>% group_by(screen_name) %>% summarize(n = n()) %>% arrange(desc(n))
hadley_fav_count %>% top_n(5, n) %>%
ggplot() +
  geom_col(mapping = aes(x = fct_reorder(screen_name, -n), y = n)) +
  scale_y_continuous(labels = comma) +
  theme(axis.title.x = element_blank()) +
  theme(axis.title.y = element_blank()) +
  labs(title = "Users Hadley Most Recently Favorited by Count",
       caption = "\nSource: Data collected via rtweet - graphic by @mjhendrickson")
```



## Upset Plot {.smaller}
``` {r upset, echo = TRUE, message = FALSE, warning = FALSE, eval = FALSE}
# Credit to @PaulCampbell91 for the borrowed code
rstaters <- c("mjhendrickson", "CMastication", "thomas_mock", "kearneymw", "LittleMissData", "AnnaHenschel")

followers_upset <- map_df(rstaters, ~ get_followers(.x, n=100000, retryonratelimit = TRUE) %>%
                          mutate(account = .x))

aRdent_followers <- unique(followers_upset$user_id)

binaries <- rstaters %>%
  map_dfc(~ ifelse(aRdent_followers %in% filter(followers_upset, account == .x)$user_id, 1, 0) %>% 
            as.data.frame)

names(binaries) <- rstaters

upset(binaries, 
      nsets = 6, 
      main.bar.color = "SteelBlue", 
      sets.bar.color = "DarkCyan", 
      sets.x.label = "Follower Count", 
      text.scale = c(rep(1.4, 5), 1), 
      order.by = "freq")
```



## Upset Plot
<center><img src = "images/plot_upsetr.png" width="750" height="419"></center>



## But wait - there's more! {.smaller}
For a full list of what's included in the data pull, go to the [Tweet Data Dictionary](https://developer.twitter.com/en/docs/tweets/data-dictionary/overview/tweet-object.html)

A few of interest:

- `created_at` = UTC time of Tweet
- `text` = content of the Tweet
- `source` = utility used to post the Tweet
- `is_retweet` = TRUE/FALSE if this was a re-tweet
- `favorite_count` = number of times the Tweet was favorited
- `retweet_count` = number of times the Tweet was re-tweeted
- `hashtags` = string of all hashtags used
- `urls_url` = string of all urls
- `mentions_screen_name` = string of screen_names mentioned
- Fields on re-tweets, geolocation, user info



## All available fields {.smaller}
``` {r dictionary, echo = TRUE, message = FALSE, warning = FALSE, eval = TRUE}
as_tibble(colnames(search_tweets("#rladies", n = 1, include_rts = FALSE)))
```



# Thank you
<img src = "images/twitter.png" width="30" height="30">  [\@mjhendrickson](https://twitter.com/mjhendrickson)

<img src = "images/linkedin.png" width="30" height="30">  [matthewjhendrickson](https://www.linkedin.com/in/matthewjhendrickson/)

<img src = "images/github.png" width="30" height="30">  [mjhendrickson](https://github.com/mjhendrickson)


This talk is freely distributed under the MIT License.  
(So is rtweet!)



## References {.smaller}
- Kearney MW (2019). "rtweet: Collecting and analyzing Twitter data." Journal of Open Source Software, 4(42), 1829. doi: 10.21105/joss.01829, R package version 0.7.0, [https://joss.theoj.org/papers/10.21105/joss.01829](https://joss.theoj.org/papers/10.21105/joss.01829).
- Rudis, B (2018). "21 Recipes for Mining Twitter with rtweet" [https://rud.is/books/21-recipes/](https://rud.is/books/21-recipes/)
- Ellis, Laura (2019). "Set Analysis: A face off between Venn diagrams and UpSet plots." [https://www.littlemissdata.com/blog/set-analysis](https://www.littlemissdata.com/blog/set-analysis)
- Kearney MW (2018). "R: Collecting and Analyzing Twitter Data." - Specifically Slide 36 on SNA. [https://github.com/mkearney/nicar_tworkshop](https://github.com/mkearney/nicar_tworkshop)
- Headshot by [Maureen Porto Photography](https://www.maureenporto.com/)
- No way gif by [https://memecrunch.com/meme/C0QML/no-way](https://memecrunch.com/meme/C0QML/no-way)
- Social media icons by [https://iconmonstr.com/](https://iconmonstr.com/)
- Twitter Data Dictionary [https://developer.twitter.com/en/docs/tweets/data-dictionary/overview/tweet-object.html](https://developer.twitter.com/en/docs/tweets/data-dictionary/overview/tweet-object.html)

